Genetic draft and valley crossing

### 2019-05-07 ms/valley_crossing.tex
further manuscript edits

Valley crossing in asexual populations is increasingly well understood \citep{weissman_2009}.
If the number of mutants produced every generation is high, as is the case in large populations with high mutation rates, a complex adaptation requiring multiple mutations that is destined to fix will eventually be generated de novo.
This is sometimes called deterministic fixation of the multiple mutant \citep{weissman_2009}.
In small populations, neutral or deleterious intermediate mutants generated by wild type individuals can drift to fixation. Further mutations necessary for the adaptation can then arise on intermediate mutant backgrounds and fix due to positive selection.
This is the sequential fixation regime \citep{weissman_2009}.
A third possibility occurs for intermediate population sizes wherein wild type individuals generate transient mutant subpopulations or ``bubbles".
These bubbles are ordinarily doomed to extinction due to drift, but additional mutations in a lucky bubble may generate a complex adaptation that sweeps to fixation; this process has been referred to as ``tunneling" \citep{iwasa_2004, weissman_2009}.
On its own, valley crossing becomes more likely as the population size increases \citep{weissman_2009}. However, recent work suggests that valley crossing relative to the fixation of a simple beneficial mutation is least likely at intermediate population sizes in which tunneling occurs \citet{ochs_2015}. This suggests that tunneling may be the most difficult mode of valley crossing. 
Recombination also affects the rate of valley crossing.
Infrequent recombination decreases valley crossing times \citep{weissman_2010}.
At high values of the recombination rate, however, individuals carrying the full complex adaptation outcross with the wild type and produce deleterious intermediates that retard the growth of the multiple mutant.
An additional effect that can heighten the rate at which valley crossing occurs is population subdivision \citep{Bitbol:Schwab:2014}: the population size within a subdivision is smaller, meaning that selection against potentially harmful intermediates is relaxed.

Previous studies of valley crossing focus primarily on populations where all individuals are equal in fitness except for the focal loci, i.e., the loci at which the individual mutations comprising the complex adaptation segregate.
This is tantamount to assuming that, if there is variation in the background fitness of the population, it is negligible.
In such populations, genetic drift governs the fate of neutral alleles, as well as the behavior of deleterious or beneficial alleles where they are rare (close to frequency zero) or common (close to frequency one).
Alleles whose dynamics are primarily determined by genetic drift can be effectively modeled by a diffusion approximation \citep{Wright:1945,Kimura:1955,Kimura:1957}, and the ancestry of a population is described well by the classic Kingman coalescent \citep{Kingman:1982}, where pairs of branches in the genealogy merge back in time until the most recent common ancestor.
When population sizes are very large or the fitness variation in the population is substantial, however, the behavior of genetic variation is governed more by genetic \emph{draft} than by genetic drift \citep{gillespie_2000, gillespie_2001, masel_2011, neher_shraiman_2011}.
Whereas drift is the effect of imperfect sampling from generation to generation, draft is the sum of hitchhiking effects due to selection on sites linked to the focal loci.
The standard technique of defining a reduced ``effective population size" fails to capture several important qualitative differences between drift and draft.
Draft is fundamentally different from drift: for example, new mutations in a drafting population can rapidly be brought to high frequency, and the offspring number variance almost diverges \citep{neher_shraiman_2011}.
Drafting populations commonly have genealogies in which more than two branches merge at once and that are better described by alternative coalescent processes such as the Bolthausen-Sznitman coalescent \citep{neher_hallatschek_2013, brunet_2007, schweinsberg_2017}.
Additionally under genetic draft, the frequency spectrum of neutral alleles is non-monotonic, with a marked uptick near frequency one--or, equivalently, a depression in intermediate frequency alleles, which are quickly swept out of the population \citep{neher_shraiman_2011,kosheleva_2013,neher_hallatschek_2013}.
Only a handful of individuals near the ``nose" of the fitness distribution are likely to persist, and they give rise to the bulk of the future population and carry linked alleles with them.
The parameter that determines whether drift or draft is more important is the product of the population size $N$ and the standard deviation in fitness $\sigma$ \citep{neher_hallatschek_2013}.

@@

The complicating factor is that, in an asexual population, the dynamics are governed almost entirely by what happens in the nose of the fitness distribution, and these dynamics can be highly stochastic.
In previous approaches, these stochastic effects were smoothed due to the presence of recombination \citep{neher_shraiman_2011, neher_shraiman_2009}: the critical nose dynamics are still important in a sexual population, but the fitness distribution is more Gaussian overall.
We therefore must focus primarily on simulation approaches, as analytical solutions for the behavior of a mutation in the nose of an asexual population are very difficult to obtain.
We show that fitness valley crossing occurs at an overall lower rate in rapidly adapting populations.
This is consistent with the fact that clonal interference and genetic background effects retard the growth of fit mutant lineages and make a given selective sweep less likely. (reword this later)
In addition, we confirm that in rapidly adapting populations, adaptive fixations of alleles are more likely to be complex adaptations involving a fitness valley than they are in populations where genetic drift is the primary force shaping genetic variation at linked sites; essentially, genetic draft maintains the linkage that makes it possible to leap across fitness landscapes rather than adapt primarily by climbing to local peaks.
These observations add to the growing intuition that complex adaptations that involve evolution across fitness valleys may not be a surprising result of the evolutionary process, but rather an expected one.

@@

Unfortunately, a number of complicating factors prevent a full analytical description of this process.
First, $\pfix$ will not depend solely on the background fitness $x$ of a lineage.
Individuals in each lineage will accumulate mutations, broadening the fitness distribution \emph{of that lineage}, so $\pfix$ will depend specifically on the background fitness of the precise individual within said lineage that gives rise to the complex adaptation.
This issue can be swept aside by assuming that bubbles do not last long enough for their own distribution of fitnesses to become very broad.
The second issue is somewhat more severe: there is no principled way to compute the Laplace transform of the bubble size distribution.
Previous approaches such as \citep{weissman_2009} and \citep{neher_shraiman_2011} have relied on branching process approximations.
\citep{neher_shraiman_2009} in particular were able to derive the asymptotics of the bubble size distribution in a facultatively sexual population: they began by deriving a master equation for the distribution of bubble sizes with a birth rate that decreases in time as the mean fitness advances.
However, they relied on a recombination model that allowed them to directly obtain the integrated Laplace transform $\int_{-\infty}^\infty p(x) \phi(x) dx$ without the need to explicitly solve the branching process model; our model being an asexual one, this option is not available to us.
\iffalse
In addition, they evaded the need to send time to infinity in the model.
In principle this is necessary to integrate over all possible bubble sizes: but in practice, because the birth rate of a lineage decreases with time, this would yield the absurd result that $\phi$ itself decreases with time.
\fi
Since it seems unlikely that analytical approaches will yield a principled way of approximating the valley crossing time in the drift case, we focus on simulation methods to estimate the valley crossing time.

@@

To analyze our model, we perform forward simulations using the custom Julia module \texttt{WaveFit} that implements a modified Wright-Fisher model.
The simulation code is available via \texttt{GitHub}.
In \texttt{WaveFit}, individuals are organized by ``clones" (sets of individuals with the same genome): each clone gives rise to a Poisson distributed number of offspring, with the mean dependent on the relative fitness $x-\bar{x}$.
The mean offspring number is further adjusted to keep the population roughly at a pre-specified carrying capacity $K$.
In addition, each individual receives a Poisson distributed number of mutations: if $L$ is the number of sites in the genome and $U$ is the (background) per-site mutation rate, then $UL$ is the mean number of mutations per individual per generation (and is a parameter in our simulations).
In effect, this is an ``infinite sites" model \citep{Kimura:1971,Watterson:1975}.
All background loci are assigned a common fitness effect $\epsilon$, which is scaled up or down each generation so as to keep the background fitness standard deviation $\sigma$ constant.
This rescaling means that the assigned initial value of $\epsilon$ turns out not to matter, even if it is negative or even if both deleterious and beneficial background mutations are permitted.
That feature of our model may seem counterintuitive, but previous work has found that populations experiencing strong background selection at many loci of weak effect (see, e.g., \citet{cvijovic_2015}) exhibit dynamics very similar to rapidly adapting populations \citep{desai_fisher_2007, neher_shraiman_2011}.

@@

For all simulations, we initialize a homogeneous wild type population (in which each individual has zero background mutations) of size $N = K$.
The precise value of $UL$ seems not to affect the valley crossing dynamics provided it is large enough for the ensemble average of the fitness distribution to be roughly Gaussian.
Accordingly, we set $UL = 0.1$, as a value less than $1$ reduces the number of operations needed for the mutation step.
We permit the population to equilibrate for $N/10$ generations.
While this is usually less than the coalescence time, it is long enough for the average number of clones to reach its approximate equilibrium value (at which the creation of new clones by mutation roughly cancels out the extinction of unfit clones).
After the population has equilibrated, we set the mutation rates at the focal loci to their pre-selected values $\mu_1$ and $\mu_2$.

We partitioned our simulations as follows.
We first studied the effect of increasing $\sigma$ in the sequential fixation, stochastic tunneling, semi-deterministic tunneling, and deterministic fixation regimes \citep{weissman_2009}.
For small populations, sequential fixation applies when $N \ll \min(1/\delta, 1/\sqrt{\mu s})$, so that a successful mutant is unlikely to arise until the intermediate has fixed, but the intermediate's fitness disadvantage is too weak for selection to inhibit it.
Tunneling becomes relevant when $N$ is too large for sequential fixation, i.e., the tunneling probability becomes more significant than the probability that the intermediate will fix: this mandates $N \gtrsim 1/\sqrt{\mu s}$.
It also requires that $N \ll 1/\mu$, meaning that only one mutant lineage is likely to be extant at any given time and, if lineages do co-occur, they are unlikely to interfere with each other or substantially affect the mean fitness.
By increasing $N$ far above the $1/\mu$ limit, multiple deleterious mutants are likely to occur every generation, reaching mutation-selection balance at low frequency and providing many opportunities for a lucky double mutant to appear; this is semi-deterministic tunneling.
At still higher population sizes ($N \gtrsim s/\mu^2$), multiple double mutants are likely to appear every generation, and their dynamics can be modeled as deterministic.

Finally, we compared the dynamics of valley crossing versus sweeping mutations.
We performed simulations with a sweeping focal beneficial mutation, with selection coefficient $s_{\mathrm{sweep}}$, and a constant influx of background beneficial mutations of weak effect as in the valley crossing simulations.
We recorded the time for a beneficial mutation to arise and reach frequency $0.5$, considering it to have fixed at that point.
We then computed the ratio of the sweep time to the valley crossing time for a complex adaptation, with valley depth $\delta$ and larger fitness advantage $s_{\mathrm{valley}}$.
To determine the effect of $\sigma$ on this ratio, we performed simulations at both high and low $\sigma$ and computed the ratio of these ratios: the result is a measure of the extent to which rapid adaptation increases the likelihood of valley crossing relative to the speed of simple sweeps.

@@

\begin{figure}[t]
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_tunnel.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\begin{subfigure}[c]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\begin{subfigure}[d]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\caption{Valley crossing times across a range of parameter values mirroring those of \citet{weissman_2009}. In figures (a) and (b), the population size is varied: increasing the size causes valley crossing to transition from tunneling to semi-deterministic fixation. In figure (c), the valley depth $\delta$ is varied: at negative $\delta$, our simulations closely match the expectation for beneficial mutations. In figure (d), the intermediate mutation rate $\mu_2$ is varied.}
\label{fig:weissman}
\end{figure}

\begin{figure}[t]
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_tunnel.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\caption{Valley crossing times in the semi-deterministic tunneling regime as a function of $\sigma$. In both figures, $s = 10^{-2}$ and $\delta = 10^{-3}$: on the left, $N = 10^5$ and $\mu = 10^{-4}$, and on the right, $N = 10^6$ and $\mu = 10^{-3}$; the right figure has a high enough value of $N\mu^2$ that fixation is almost deterministic. Orange lines are averages, boxes are interquartile ranges, and end caps are overall ranges. Each box represents about $100$ simulation runs.}
\label{fig:deterministic}
\end{figure}

\begin{figure}[t]
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/seq_fix.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/neut_tunnel.pdf}
\end{subfigure}
\caption{Valley crossing times in the sequential fixation (left) and tunneling (right) regimes as a function of $\sigma$. In both figures, $s = 10^{-2}$ and $\delta = 10^{-3}$: on the left, $N = 10^4$ and $\mu = 3\times 10^{-5}$, and on the right, $N = 10^5$ and $\mu = 3\times 10^{-6}$. Labeling is as in figure \ref{fig:deterministic}.}
\label{fig:tunneling}
\end{figure}

Here we present the key results of our simulation analysis.
In general, we expect valley crossing to require more time in rapidly adapting populations (large $N \sigma$) experiencing genetic draft; bubbles are doomed to extinction more quickly (lineages experience larger fluctuations in size and decline in frequency as the mean fitness advances) and beneficial lineages are less likely to fix, especially when valleys are shallow.
Figures \ref{fig:deterministic} and \ref{fig:tunneling} confirm these expectations: the overall rate of valley crossing is generally slowed as $\sigma$ increases.
At higher $\sigma$, the effect of the genetic background in which an mutation finds itself becomes more significant in determining the fate of that mutation than the fitness effect of the mutation itself.
This genetic draft buffets the frequency of mutations so that these mutations behave as though they are more neutral than they really are: in effect, the fitness landscape is ``flattened".
This depresses the fixation probability of the double mutant and other complex adaptations, which appears to account for the majority of the change in fixation time.
In the case of stochastic tunneling, the overall size distribution of bubble sizes is reduced as well: the distribution $P(w)$ scales not as $w^{-3/2}$ but as $w^{-2}$ \citep{neher_shraiman_2011}, so there are fewer chances for the complex adaptation to arise on a mutant background and fix.

\begin{figure}[t]
\includegraphics[width=0.4\textwidth]{Figures/var_sigma_delta.pdf}
\caption{Valley crossing times for varied $\sigma$ and $\delta$. In these simulations, $N = 10^5$, $\mu = 3 \times 10^{-6}$, and $s = 10^{-2}$. Note the lack of dependence on $\delta$ at high $\sigma$.}
\label{fig:sigma_delta}
\end{figure}

@@

\begin{figure}[t]
\begin{subfigure}[b]{0.475\textwidth}
  \includegraphics[width=\textwidth]{Figures/tau_compare_crop.pdf}
  \caption{Valley crossing times on a $\log_{10}$ scale}
  \label{fig:tau}
\end{subfigure}
\begin{subfigure}[b]{0.475\textwidth}
  \includegraphics[width=\textwidth]{Figures/alpha_ratios_crop.pdf}
  \caption{Ratio $\alpha = \tau_{\mathrm{valley}}/\tau_{\mathrm{sweep}}$ on a $\log_{10}$ scale}
  \label{fig:ratios}
\end{subfigure}
\caption{Behavior of valley crossing times for varied $\delta$, $N$, and $\sigma$. In all cases, $s_{\mathrm{sweep}} = 0.01$, $s_{\mathrm{valley}} = 0.1$, and $\mu = 10^{-5}$. Panel (a) shows that valley crossing times $\tau$ increase with $\delta$ and $\sigma$ but decrease with $N$, as predicted. Panel (b) shows the ratio $\alpha = \tau_{\mathrm{valley}}/\tau_{\mathrm{sweep}}$ as a function of $\delta$, $N$, and $\sigma$.}
\label{fig:tau-ratios}
\end{figure}

>>>>

Valley crossing in asexual populations is increasingly well understood \citep{weissman_2009}.
If the number of mutants produced every generation is high, as is the case in large populations with high mutation rates, a complex adaptation requiring multiple mutations that is destined to fix will eventually be generated \emph{de novo}.
This is sometimes called deterministic fixation of the multiple mutant \citep{weissman_2009}.
In small populations, neutral or deleterious intermediate mutants generated by wild type individuals can drift to fixation. Further mutations necessary for the adaptation can then arise on intermediate mutant backgrounds and fix due to positive selection.
This is the sequential fixation regime \citep{weissman_2009}.
A third possibility occurs for intermediate population sizes wherein wild type individuals generate transient mutant subpopulations or ``bubbles".
These bubbles are ordinarily doomed to extinction due to drift, but additional mutations in a lucky bubble may generate a complex adaptation that sweeps to fixation; this process has been referred to as ``tunneling" \citep{iwasa_2004, weissman_2009}.
On its own, valley crossing becomes more likely as the population size increases \citep{weissman_2009}. However, recent work suggests that valley crossing relative to the fixation of a simple beneficial mutation is least likely at intermediate population sizes in which tunneling occurs \citet{ochs_2015}. This suggests that tunneling may be the most difficult mode of valley crossing. 
Recombination also affects the rate of valley crossing.
Infrequent recombination decreases valley crossing times \citep{weissman_2010}.
At high values of the recombination rate, however, individuals carrying the full complex adaptation outcross with the wild type and produce deleterious intermediates that retard the growth of the multiple mutant.
An additional effect that can heighten the rate at which valley crossing occurs is population subdivision \citep{Bitbol:Schwab:2014}: the population size within a subdivision is smaller, meaning that selection against potentially harmful intermediates is relaxed.

Previous studies of valley crossing focus primarily on populations where all individuals are equal in fitness except for the focal loci, i.e., the loci at which the individual mutations comprising the complex adaptation segregate.
This is tantamount to assuming that, if there is variation in the background fitness of the population, it is negligible.
In such populations, genetic drift governs the fate of neutral alleles, as well as the behavior of deleterious or beneficial alleles where they are rare (close to frequency zero) or common (close to frequency one).
Alleles whose dynamics are primarily determined by genetic drift can be effectively modeled by a diffusion approximation \citep{Wright:1945,Kimura:1955,Kimura:1957}, and the ancestry of a population in the absence of selection is described well by the classic Kingman coalescent \citep{Kingman:1982}, where pairs of branches in the genealogy merge back in time until the most recent common ancestor.
When population sizes are very large or the fitness variation in the population is substantial, however, the behavior of genetic variation is governed more by genetic \emph{draft} than by genetic drift \citep{gillespie_2000, gillespie_2001, masel_2011, neher_shraiman_2011}.
Whereas drift is the effect of imperfect sampling from generation to generation, draft is the sum of hitchhiking effects due to selection on sites linked to the focal loci.
The standard technique of defining a reduced ``effective population size" fails to capture several fundamental qualitative differences between drift and draft.
For example, new mutations in a drafting population can rapidly be brought to high frequency, and the variance in offspring number almost diverges \citep{neher_shraiman_2011}.
Drafting populations commonly have genealogies in which more than two branches merge at once (or almost at once) and that are better described by alternative coalescent processes such as the Bolthausen-Sznitman coalescent \citep{neher_hallatschek_2013, brunet_2007, schweinsberg_2017}.
Additionally under genetic draft, the frequency spectrum of neutral alleles is non-monotonic, with a marked uptick near frequency one--or, equivalently, a depression in intermediate frequency alleles, which are quickly swept out of the population \citep{neher_shraiman_2011,kosheleva_2013,neher_hallatschek_2013}.
Only a handful of individuals near the ``nose" of the fitness distribution are likely to persist, and they give rise to the bulk of the future population and carry linked alleles with them.
The parameter that determines whether drift or draft is more important is the product of the population size $N$ and the standard deviation in fitness $\sigma$ \citep{neher_hallatschek_2013}.

@@

The complicating factor is that, in an asexual population, the dynamics are governed almost entirely by what happens in the nose of the fitness distribution, and these dynamics can be highly stochastic.
In previous approaches, these stochastic effects were smoothed due to the presence of recombination \citep{neher_shraiman_2011, neher_shraiman_2009}: the critical nose dynamics are still important in a sexual population, but the fitness distribution is more Gaussian overall.
We therefore must focus primarily on simulation approaches, as analytical solutions for the behavior of a mutation in the nose of an asexual population are very difficult to obtain.
We expand on the work of \citet{neher_shraiman_2011} by considering not only stochastic tunneling but also sequential fixation; in fact, we find that, over a significant chunk of parameter space, stochastic tunneling effectively \emph{does not happen} in drafting populations, and sequential fixation appears to be the dominant mode of valley crossing.
We also show that fitness valley crossing occurs at an overall lower rate in rapidly adapting populations.
This is consistent with the fact that clonal interference and genetic background effects retard the growth of fit mutant lineages, slowing the expansion of a beneficial mutation.
In addition, we confirm that in rapidly adapting populations, adaptive fixations of alleles are more likely to be complex adaptations involving a fitness valley than they are in populations where genetic drift is the primary force shaping genetic variation at linked sites; essentially, genetic draft maintains the linkage that makes it possible to leap across fitness landscapes rather than adapt primarily by climbing to local peaks.
These observations add to the growing intuition that complex adaptations that involve evolution across fitness valleys may not be a surprising result of the evolutionary process, but rather an expected one.

@@

Unfortunately, a number of complicating factors prevent a full analytical description of this process.
First, $\pfix$ will not depend solely on the background fitness $x$ of a lineage.
Individuals in each lineage will accumulate mutations, broadening the fitness distribution \emph{of that lineage}, so $\pfix$ will depend specifically on the background fitness of the precise individual within said lineage that gives rise to the complex adaptation.
This issue can be swept aside by assuming that bubbles do not last long enough for their own distribution of fitnesses to become very broad.
The second issue is somewhat more severe: there is no principled way to compute the Laplace transform of the bubble size distribution.
Previous approaches such as \citet{weissman_2009} and \citet{neher_shraiman_2011} have relied on branching process approximations.
The latter in particular were able to derive the asymptotics of the bubble size distribution in a facultatively sexual population: they began by deriving a master equation for the distribution of bubble sizes with a birth rate that decreases in time as the mean fitness advances.
However, they relied on a recombination model that allowed them to directly obtain the integrated Laplace transform $\int_{-\infty}^\infty p(x) \phi(x) dx$ without the need to explicitly solve the branching process model; our model being an asexual one, this option is not available to us.
\iffalse
In addition, they evaded the need to send time to infinity in the model.
In principle this is necessary to integrate over all possible bubble sizes: but in practice, because the birth rate of a lineage decreases with time, this would yield the absurd result that $\phi$ itself decreases with time.
\fi
Since it seems unlikely that analytical approaches will yield a principled way of approximating the valley crossing time in the drift case, we focus on simulation methods to estimate the valley crossing time.

@@

To analyze our model, we perform forward simulations using the custom Julia module \texttt{WaveFit}, which implements a modified Wright-Fisher model.
The simulation code is available via \texttt{GitHub}.
In \texttt{WaveFit}, individuals are organized by ``clones" (sets of individuals with the same genome): each clone gives rise to a Poisson distributed number of offspring, with the mean dependent on the relative fitness $x-\bar{x}$.
The mean offspring number is further adjusted to keep the population roughly at a pre-specified carrying capacity $K$.
In addition, each individual receives a Poisson distributed number of mutations: if $L$ is the number of sites in the genome and $U$ is the (background) per-site mutation rate, then $UL$ is the mean number of mutations per individual per generation (and is a parameter in our simulations).
In effect, this is an ``infinite sites" model \citep{Kimura:1971,Watterson:1975}.
All background loci are assigned a common fitness effect $\epsilon$, which is scaled up or down each generation so as to keep the background fitness standard deviation $\sigma$ constant.
This rescaling means that the assigned initial value of $\epsilon$ turns out not to matter, even if it is negative or even if both deleterious and beneficial background mutations are permitted.
That feature of our model may seem counterintuitive, but previous work has found that populations experiencing strong background selection at many loci of weak effect (see, e.g., \citet{cvijovic_2015}) exhibit dynamics very similar to rapidly adapting populations \citep{desai_fisher_2007, neher_shraiman_2011}.

@@

For all simulations, we initialize a homogeneous wild type population (in which each individual has zero background mutations) of size $N = K$, with $\mu_1 = \mu_2 = 0$.
The precise value of $UL$ seems not to affect the valley crossing dynamics provided it is large enough for the ensemble average of the fitness distribution to be roughly Gaussian (see appendix).
Accordingly, we set $UL = 1.0$.
We then permit the population to equilibrate for $N/10$ generations.
While this is usually less than the coalescence time, it is long enough for the average number of clones to reach its approximate equilibrium value (at which the creation of new clones by mutation roughly cancels out the extinction of unfit clones); see appendix.
After the population has equilibrated, we set the mutation rates at the focal loci to their pre-selected values $\mu_1$ and $\mu_2$.

We partitioned our simulations as follows.
We first considered the parameter regimes modeled in \citet{weissman_2009}, which were previously studied solely in the presence of genetic drift, and introduced genetic draft by turning up $\sigma$.
In the drift case, the dynamics are as follows: we comment on important differences for the draft case in the results section.
For small populations, sequential fixation applies when $N \ll \min(1/\delta, 1/\sqrt{\mu s})$, so that a successful mutant is unlikely to arise until the intermediate has fixed, but the intermediate's fitness disadvantage is too weak for selection to inhibit it.
Tunneling becomes relevant when $N$ is too large for sequential fixation, i.e., the tunneling probability becomes more significant than the probability that the intermediate will fix: this mandates $N \gtrsim 1/\sqrt{\mu s}$.
It also requires that $N \ll 1/\mu$, meaning that only one mutant lineage is likely to be extant at any given time and, if lineages do co-occur, they are unlikely to interfere with each other or substantially affect the mean fitness.
By increasing $N$ far above the $1/\mu$ limit, multiple deleterious mutants are likely to occur every generation, reaching mutation-selection balance at low frequency and providing many opportunities for a lucky double mutant to appear; this is semi-deterministic tunneling.
At still higher population sizes ($N \gtrsim s/\mu^2$), multiple double mutants are likely to appear every generation, and their dynamics can be modeled as deterministic.

Next, we directly examined the effects of varying both $\sigma$ and $\delta$ on the valley crossing time.

Finally, we compared the dynamics of valley crossing versus sweeping mutations.
We performed simulations with a sweeping focal beneficial mutation, with selection coefficient $s_{\mathrm{sweep}}$, and a constant influx of background beneficial mutations of weak effect as in the valley crossing simulations.
We recorded the time for a beneficial mutation to arise and reach frequency $0.5$, considering it to have fixed at that point.
We then computed the ratio of the sweep time to the valley crossing time for a complex adaptation, with valley depth $\delta$ and larger fitness advantage $s_{\mathrm{valley}}$.
To determine the effect of $\sigma$ on this ratio, we performed simulations at both high and low $\sigma$ and computed the ratio of these ratios: the result is a measure of the extent to which rapid adaptation facilitates valley crossing rather than simple selective sweeps.

@@

\begin{figure*}[t]
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5a.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5b.pdf}
\end{subfigure}
\begin{subfigure}[c]{0.5\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5c.pdf}
\end{subfigure}
\begin{subfigure}[d]{0.5\textwidth}
\includegraphics[width=\textwidth]{Figures/julia_weissman_5d.pdf}
\end{subfigure}
\caption{Valley crossing times across a range of parameter values mirroring those of \citet{weissman_2009}. In figures (a) and (b), the population size is varied: increasing the size causes valley crossing to transition from tunneling to semi-deterministic fixation. In figure (c), the valley depth $\delta$ is varied: at negative $\delta$, our simulations closely match the expectation for beneficial mutations. In figure (d), the intermediate mutation rate $\mu_2$ is varied. Dotted lines represent the theoretical predictions from \citet{weissman_2009}.}
\label{fig:weissman}
\end{figure*}

\iffalse
\begin{figure}[t]
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_tunnel.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/det_fix.pdf}
\end{subfigure}
\caption{Valley crossing times in the semi-deterministic tunneling regime as a function of $\sigma$. In both figures, $s = 10^{-2}$ and $\delta = 10^{-3}$: on the left, $N = 10^5$ and $\mu = 10^{-4}$, and on the right, $N = 10^6$ and $\mu = 10^{-3}$; the right figure has a high enough value of $N\mu^2$ that fixation is almost deterministic. Orange lines are averages, boxes are interquartile ranges, and end caps are overall ranges. Each box represents about $100$ simulation runs.}
\label{fig:deterministic}
\end{figure}

\begin{figure}[t]
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/seq_fix.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.4\textwidth}
\includegraphics[width=\textwidth]{Figures/neut_tunnel.pdf}
\end{subfigure}
\caption{Valley crossing times in the sequential fixation (left) and tunneling (right) regimes as a function of $\sigma$. In both figures, $s = 10^{-2}$ and $\delta = 10^{-3}$: on the left, $N = 10^4$ and $\mu = 3\times 10^{-5}$, and on the right, $N = 10^5$ and $\mu = 3\times 10^{-6}$. Labeling is as in figure \ref{fig:deterministic}.}
\label{fig:tunneling}
\end{figure}
\fi

Here we present the key results of our simulation analysis.
Figure \ref{fig:weissman} summarizes the major differences between the drift (low $\sigma$) and draft (high $\sigma$) regimes in valley crossing.
Note that the theoretical predictions differ slightly from those of \citet{weissman_2009} due to their use of a Moran model, compared to our Wright-Fisher model: this is comparable to an order of 2 difference in the population size.
The general trend is that draft causes valley crossing to take more time but be less sensitive to the underlying parameter values.
In figure (a), tunneling at low $\sigma$ occurs via neutral sequential fixation at low $N$ before transitioning into neutral tunneling and semi-deterministic tunneling at higher $N$; these are visible as minor kinks in the theoretical predictions (dotted line).
At low $\sigma$, our results agree with those predictions, but as $\sigma$ increases, the dependence on $N$ is lessened: at higher $\sigma$, the population effectively behaves as though there is a very small $N_e$, resulting in no dependence on $N$.
These trends are also visible in figure (c) where, in the drift case, the population transitions from neutral to deleterious tunneling, and figure (d), in which deleterious tunneling gives way to neutral tunneling as the mutation rate $\mu_2$ increases.

One remarkable difference between drift and draft can be seen in figure (b).
At $\sigma = 10^{-8}$, crossing proceeds via deleterious sequential fixation at low $N$, then transitions into deleterious tunneling around $N = 10^3$ and semi-deterministic tunneling at higher $N$.
The increase in crossing time with $N$ is a hallmark of deleterious sequential fixation: higher $N$ means the deleterious intermediate is more ``visible" to selection, slowing its overall growth.
Once the population becomes large enough that fixation of the intermediate is unlikely, raising $N$ increases the number of single mutant ``bubbles" that are generated, resulting in the $\tau \sim N^{-1}$ crossing time dependence.
As $\sigma$ approaches even the moderate value $10^{-2}$, the crossing time curve flattens significantly; it still positively depends on $N$ but never approaches the peak crossing time $\approx 10^7$ generations in the $\sigma = 10^{-8}$ case.
The behavior in these two cases is qualitatively different: the intermediate $\sigma$ curve's dependence on $N$ cannot be reduced to the drift case simply by defining a reduced $N_e$.
In effect, from the point of view of the focal mutations, draft flattens the fitness landscape somewhat but not completely.
At higher values of $\sigma$, $\tau$ is again mostly independent of $N$, an indicator that sequential fixation is the dominant mode of crossing.

\iffalse
In general, we expect valley crossing to require more time in rapidly adapting populations (large $N \sigma$) experiencing genetic draft; bubbles are doomed to extinction more quickly (lineages experience larger fluctuations in size and decline in frequency as the mean fitness advances) and beneficial lineages are less likely to fix, especially when valleys are shallow.
Figures \ref{fig:deterministic} and \ref{fig:tunneling} confirm these expectations: the overall rate of valley crossing is generally slowed as $\sigma$ increases.
At higher $\sigma$, the effect of the genetic background in which an mutation finds itself becomes more significant in determining the fate of that mutation than the fitness effect of the mutation itself.
This genetic draft buffets the frequency of mutations so that these mutations behave as though they are more neutral than they really are: in effect, the fitness landscape is ``flattened".
This depresses the fixation probability of the double mutant and other complex adaptations, which appears to account for the majority of the change in fixation time.
In the case of stochastic tunneling, the overall size distribution of bubble sizes is reduced as well: the distribution $P(w)$ scales not as $w^{-3/2}$ but as $w^{-2}$ \citep{neher_shraiman_2011}, so there are fewer chances for the complex adaptation to arise on a mutant background and fix.
\fi

\begin{figure}[t]
\includegraphics[width=0.4\textwidth]{Figures/julia_sigma_delta.pdf}
\caption{Valley crossing times for varied $\sigma$ and $\delta$. In these simulations, $N = 10^5$, $\mu_1 = 3 \times 10^{-6}$, and $\mu_2 = 10^{-4}$. Note the lack of dependence on $\delta$ at high $\sigma$.}
\label{fig:sigma_delta}
\end{figure}

@@

\begin{figure*}[t]
\begin{subfigure}[b]{0.5\textwidth}
  \includegraphics[width=\textwidth]{Figures/tau_lowsig.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
  \includegraphics[width=\textwidth]{Figures/tau_highsig.pdf}
\end{subfigure}
\label{fig:tau}

  \caption{Valley crossing times on a $\log_{10}$ scale}
\begin{subfigure}[b]{0.5\textwidth}
  \includegraphics[width=\textwidth]{Figures/ratio_lowsig.pdf}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
  \includegraphics[width=\textwidth]{Figures/ratio_highsig.pdf}
\end{subfigure}
\caption{Behavior of valley crossing times for varied $\delta$, $N$, and $\sigma$. In all cases, $s_{\mathrm{sweep}} = 0.01$, $s_{\mathrm{valley}} = 0.1$, $\mu_1 = 3 \times 10^{-6}$, and $\mu_2 = 10^{-4}$. Panel (a) shows that valley crossing times $\tau$ increase with $\delta$ and $\sigma$ but decrease with $N$, as predicted. Panel (b) shows the ratio $\alpha = \tau_{\mathrm{valley}}/\tau_{\mathrm{sweep}}$ as a function of $\delta$, $N$, and $\sigma$.}
\label{fig:tau-ratios}
\end{figure*}
